---
title: "Data Science Capstone"
author: "Ismail Che Ani"
date: "September 13, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r , warning=FALSE, message = FALSE, echo=TRUE}
install.packages("tm", repos = "http://cran.us.r-project.org")
## Installing package into '/Users/victoriadotta/Library/R/3.3/library'
## (as 'lib' is unspecified)
## 
## The downloaded binary packages are in
##  /var/folders/st/zr32ms157s336zw_znx_l5140000gn/T//Rtmpq3qAbc/downloaded_packages
library(tm)
```
```{r setup, include=FALSE}
## Loading required package: NLP
install.packages("wordcloud", repos = "http://cran.us.r-project.org")
## Installing package into '/Users/victoriadotta/Library/R/3.3/library'
## (as 'lib' is unspecified)
## 
## The downloaded binary packages are in
##  /var/folders/st/zr32ms157s336zw_znx_l5140000gn/T//Rtmpq3qAbc/downloaded_packages
library(wordcloud)
## Loading required package: RColorBrewer
install.packages("stringi", repos = "http://cran.us.r-project.org")
## Installing package into '/Users/victoriadotta/Library/R/3.3/library'
## (as 'lib' is unspecified)
## 
## The downloaded binary packages are in
##  /var/folders/st/zr32ms157s336zw_znx_l5140000gn/T//Rtmpq3qAbc/downloaded_packages
library(stringi)
```
## Task 0: Understanding the Problem
Tasks to accomplish:

What data we have.

What are the standard tools and models used for that type of data.

The data is from a corpus called HC Corpora (www.corpora.heliohost.org). See the readme file at http://www.corpora.heliohost.org/aboutcorpus.html for details on the corpora available. The files have been language filtered but may still contain some foreign text.

We are going to use the english dataset which contains three files.

##0.1 Loading the data

```{r , warning=FALSE, message = FALSE, echo=TRUE}
setwd("D:/OneDrive/Public/Data Science/Coursera/Capstone/")
blogs <- readLines("en_US.blogs.txt", encoding = "UTF-8", skipNul=TRUE)
news <- readLines("en_US.news.txt", encoding = "UTF-8", skipNul = TRUE)
twitter <- readLines("en_US.twitter.txt", encoding = "UTF-8", skipNul = TRUE)
```
```{r , warning=FALSE, message = FALSE, echo=TRUE}

```
